### 2025-05-05 chapters/4-substructure.tex
Pre-correction version of thesis

    \caption[Safety and continuity properties of the \PIRANHA{} and Soft Drop grooming algorithms.]{
    Safety and continuity properties of the \PIRANHA{} and Soft Drop grooming algorithms studied in this chapter.
    %
    The mark \cmark~indicates that a groomer satisfies the associated condition, while the mark \xmark~indicates that it does not.
    %
    The mark \raisebox{0.15 em}{\danger} indicates that a groomer does not satisfy the condition, but only in a boundary region of phase space that is exponentially suppressed at leading order in perturbative QCD.
    %
    The mark \raisebox{0.15 em}{\danger} indicates that a groomer does not satisfy the condition, but only in a boundary region of phase space that is exponentially suppressed at leading order in perturbative QCD.\footnotemark
    }
    \label{tab:groomerlist}
\end{table}

@@

%==============================================
\section[\textbf{P}ileup and \textbf{I}nfrared \textbf{R}adiation \textbf{A}n\textbf{n}i\textbf{h}il\textbf{a}tion (\textsc{Piranha}):\\A Paradigm for Continuous Jet Grooming]{\textbf{P}ileup and \textbf{I}nfrared \textbf{R}adiation \textbf{A}n\textbf{n}i\textbf{h}il\textbf{a}tion (PIRANHA):\\A Paradigm for Continuous Jet Grooming}
%==============================================
\label{sec:piranha}
% TODO: Make sure this markright is in the right place even at the beginning of the section
\markright{\thechapter.\thesection{} \textsc{Piranha}: A Paradigm for Continuous Jet Grooming}

@@

%=====================================
% Mass Resolution:
%=====================================
\subsection{Mass Extraction with Grooming: \texorpdfstring{\(W\)}{W} and Top Jets at the LHC}
\label{sec:pira-mass}
% TODO: add gls mass extraction to PIRANHA mass res section

    As a final case study, we explore the use of \PIRANHA{} grooming strategies in the tagging of boosted objects, another useful application of traditional grooming strategies such as Soft Drop.
   %
   We must once again overcome the subtlety that \PIRANHA{} grooming strategies remove energy proportional to the grooming parameter (such as \(\zcut\) in Recursive Safe Subtraction).
   %
   To successfully tag boosted objects decaying into jets, we must tune the grooming parameter for \PIRANHA{} strategies more than we would for Soft Drop.
   %
   Nonetheless, we examine the mass resolution of \PIRANHA{}-groomed jets from the decays of boosted \(W\) bosons and top quarks and discover that \PIRANHA{} may offer some advantages that are complementary to those of traditional grooming strategies.
   %
   The tagging procedure we implement in the following discussion is quite simple, however, and the applications of \PIRANHA{} in boosted object tagging should be subjected to a more complete analysis.

@@

% TODO: where do I put SECTION ACKNOWLEDGMENTS?
%\acknowledgments

@@

\makeprob{Constant-Cutoff Substructure at MLL}{constant-cutoff-mll}{
    Find the radiator for \glslink{constant-cutoff}{constant-cutoff-groomed} angularities at MLL.
    %
    Include running coupling effects and freeze the coupling at a non-perturbative scale \(\mu_\text{NP}\).
    %
    Write the associated MLL cumulative distribution and probability distribution in closed form.
}

\makeprob{Constant-Subtraction Substructure at MLL}{constant-subtraction-mll}{
    Find the radiator for \glslink{constant-subtraction}{constant-subtraction-groomed} angularities at MLL.
    %
    Include running coupling effects and freeze the coupling at a non-perturbative scale \(\mu_\text{NP}\).
    %
    Write the associated MLL cumulative distribution and probability distribution in closed form.
}

\makeprob{\(\star\) Multiplicity of Simple Groomed Jets \(\star\)}{simple-groomed-multiplicity}{
    Find the multiplicity distribution for constant-cutoff-groomed jets using the principles of the partonic cascade.
    %
    You may wish to refer to the solution of \Prob{subjet-multiplicity}.
}

>>>>

    \caption[Safety and continuity properties of the \PIRANHA{} and Soft Drop grooming algorithms.]{
    Safety and continuity properties of the \PIRANHA{} and Soft Drop grooming algorithms studied in this chapter.
    %
    The mark \cmark~indicates that a groomer satisfies the associated condition, while the mark \xmark~indicates that it does not.
    %
    The mark \raisebox{0.15 em}{\danger} indicates that a groomer does not satisfy the condition, but only in a boundary region of phase space that is exponentially suppressed at leading order in perturbative QCD.
    }
    \label{tab:groomerlist}
\end{table}

@@

%==============================================
\section[\textbf{P}ileup and \textbf{I}nfrared \textbf{R}adiation \textbf{A}n\textbf{n}i\textbf{h}il\textbf{a}tion (\textsc{Piranha}):\\A Paradigm for Continuous Jet Grooming]{\textbf{P}ileup and \textbf{I}nfrared \textbf{R}adiation \textbf{A}n\textbf{n}i\textbf{h}il\textbf{a}tion (PIRANHA):\\A Paradigm for Continuous Jet Grooming}
%==============================================
\label{sec:piranha}
\markright{\thechapter.\thesection{} \textsc{Piranha}: A Paradigm for Continuous Jet Grooming}

@@

%=====================================
% Mass Resolution:
%=====================================
\subsection{Mass Extraction with Grooming: \texorpdfstring{\(W\)}{W} and Top Jets at the LHC}
\label{sec:pira-mass}
    As a final case study, we explore the use of \PIRANHA{} grooming strategies in the tagging of boosted objects, another useful application of traditional grooming strategies such as Soft Drop.
   %
   We must once again overcome the subtlety that \PIRANHA{} grooming strategies remove energy proportional to the grooming parameter (such as \(\zcut\) in Recursive Safe Subtraction).
   %
   To successfully tag boosted objects decaying into jets, we must tune the grooming parameter for \PIRANHA{} strategies more than we would for Soft Drop.
   %
   Nonetheless, we examine the mass resolution of \PIRANHA{}-groomed jets from the decays of boosted \(W\) bosons and top quarks and discover that \PIRANHA{} may offer some advantages that are complementary to those of traditional grooming strategies.
   %
   The tagging procedure we implement in the following discussion is quite simple, however, and the applications of \PIRANHA{} in boosted object tagging should be subjected to a more complete analysis.

@@

%\acknowledgments

@@

    % TODO
\makeprob{Constant-Cutoff Substructure at MLL}{constant-cutoff-mll}{
    Find the radiator for \glslink{constant-cutoff}{constant-cutoff-groomed} angularities at MLL.
    %
    Include running coupling effects and freeze the coupling at a non-perturbative scale \(\mu_\text{NP}\).
    %
    Write the associated MLL cumulative distribution and probability distribution in closed form.
}

    % TODO
\makeprob{Constant-Subtraction Substructure at MLL}{constant-subtraction-mll}{
    Find the radiator for \glslink{constant-subtraction}{constant-subtraction-groomed} angularities at MLL.
    %
    Include running coupling effects and freeze the coupling at a non-perturbative scale \(\mu_\text{NP}\).
    %
    Write the associated MLL cumulative distribution and probability distribution in closed form.
}

%\makeprob{\(\star\) Multiplicity of Groomed Subjets \(\star\)}{simple-groomed-multiplicity}{
%    Using the leading-logarithmic principles of the partonic cascade, find the multiplicity distribution of subjets of radius \rsub{} within a jet of radius \Rjet{} that is groomed using the constant-cutoff algorithm.

%    \texttt{(Hint:)}
%    %
%    You may wish to refer to the solution of \Prob{subjet-multiplicity}.
%}

### 2025-05-05 chapters/5-event_shapes.tex
Pre-correction version of thesis

Importantly, we have left the sum on \(i,j\) a little bit vague.
%
What are we summing over?
%
The answer is that if we \(i\) and \(j\) to index particles, so that the \gls{ewoc} captures pair-wise particle correlations, we quickly run into trouble.
%
In particular, collinear safety requires that particle-level \glspl{ewoc} can \textit{only} encode angular correlations within a jet:
%
\(\mathcal{O}_{ij} = \theta_{ij}\) or \(\mathcal{O}_{ij} = \Delta R_{ij}\).
%
In order to obtain \glslink{irc-safety}{IRC safe} observable, we must either study only \textit{angular correlations} or \textit{give up on particle-level correlations} in favor of correlations between collective degrees of freedom.%
\footnote{
    We use subjets as proxies for collective degrees of freedom so that our energy-weighted correlations are not fully continuous \glslink{continuity}{continuous} (the number of subjets in a jet, and the discrete yes/no choice of whether a particle belongs to a given subjet, are both discontinuous).
    %
    The only fully continuous substructure observables of this chapter are the EEC and its multi-particle generalizations.
    %
    Nonetheless, the study of \glspl{ewoc} will lead to results which are \textit{more} robust to hadronization.
}

@@

Among jet substructure observables, the Energy-Energy Correlator (EEC) \cite{Basham:1978bw,Basham:1978zq,Basham:1979gh} is particularly noteworthy:
%
it is conceptually simple, insensitive to low-energy radiation, and naturally separates physics at different scales.
%
Physically, the EEC measures the mean correlation in the energy passing through any two detectors separated by an angle \(\Delta\), averaged over many particle collisions:
%
\(
    \text{EEC}(\Delta)
    \propto
    \left\langle
        E_\text{det.\.1} \,\,\, E_\text{det.\.2}
    \right\rangle
\).%
\footnote{
    More concretely, this means that the contribution of each particle in an event to the EEC is weighted by the particle's energy,
    %
    as in \Eq{intro_ewoc_def} and, shorty, \Def{eec}.
    %
    This energy weighting mitigates the effects of the low-energy physics of QCD that complicates the theoretical description of other jet substructure observables.
}
%
From the birth of the EEC to its use in recent collider studies, the simplicity of the EEC has enabled highly accurate perturbative computations \cite{Clay:1995sd,Glover:1994vz,Kramer:1996qr,DelDuca:2016csb,Gituliar:2017umx,Dixon:2018tpg,Dixon:2018qgp,Henn:2019gkr,Luo:2019nig,Gao:2020vyx,Neill:2022lqx,Lee:2023npz,Kramer:1995qh,deFlorian:2004mp,Banfi:2002vw,Tulipant:2017ybb,Moult:2018jzp,Korchemsky:2019nzm,Dixon:2019uzg,Gao:2019ojf,Luo:2019hmp,Luo:2019bmw,Moult:2019vou,Li:2020bub,Ebert:2020sfi,Li:2021txc,Duhr:2022yyp,Chen:2023zlx,Gao:2023ivm} and has facilitated measurements of the strong coupling constant \(\alpha_s\) \cite{Martin:1986uq,DELPHI:1990sof,SLD:1994yoe,ATLAS:2015yaa,ATLAS:2017qir,dEnterria:2018cye,Kardos:2018kqj,Ali:2020ksn,dEnterria:2022hzv,ATLAS:2023tgo} including the most precise jet substructure determinations of \(\alpha_s\) to date \cite{CMS:2024mlf}.
%
The EEC has been investigated as a probe of jets traveling through the quark-gluon plasma (QGP) produced in heavy-ion collisions \cite{Lokhtin:2004tx,Lokhtin:2006dp,Andres:2022ovj,Barata:2023zqg,Andres:2023xwr,Yang:2023dwc,Barata:2023vnl,Barata:2023bhh,Barata:2024nqo} and of correlations in nuclear physics \cite{Karapetyan:2019fst,Liu:2022wop,Liu:2023aqb,Kang:2023gvg,Cao:2023oef}.
%
The study of the EEC has also shed light on the structure of non-perturbative QCD \cite{Nason:1995np,Korchemsky:1997sy,Korchemsky:1999kt,Dokshitzer:1999sh,Chen:2020vvp,Jaarsma:2023ell,Schindler:2023cww,Lee:2024esz}
and of quantum field theory in general \cite{Richards:1983sr,Sveshnikov:1995vi,Hofman:2008ar,Hatta:2012kn,Belitsky:2013bja,Belitsky:2013ofa,Belitsky:2013xxa,Belitsky:2014zha,Korchemsky:2015ssa,Goncalves:2014ffa,Farnsworth:2015hum,Hofman:2016awc,Kravchuk:2018htv,Kologlu:2019mfz,Chang:2020qpj,Korchemsky:2021htm,Caron-Huot:2022eqs,Chen:2023wah,Chicherin:2023gxt,Chicherin:2024ifn}.
%
It also been put forward as a way to search for new physics in hadronic decays of electroweak bosons~\cite{Ricci:2022htc}.
%
Finally, while most jet substructure observables are dominated by contributions from two-particle correlations, the EEC can be naturally extended to multi-point energy correlators which can characterize correlations between an arbitrary number of particles.
%
Recent theoretical work suggests that multi-point generalizations of the EEC may be used to extract the mass of the top quark at the LHC \cite{Procura:2022fid,Holguin:2022epo,Holguin:2023bjf,Pathak:2023tmy,Xiao:2024rol,Holguin:2024tkz}.

Energy correlators have also been used in studies of gluon saturation and nuclear tomography  \cite{Karapetyan:2019fst,Liu:2022wop,Liu:2023aqb,Kang:2023gvg,Cao:2023oef}.
%
Further, energy correlators have produced the most precise jet substructure measurement of the strong coupling constant to date~\cite{CMS:2024mlf}.

@@

\remark{}{
    As discussed in , The EEC is \glslink{irc-safety}{IRC safe} (as we discuss further in \Sec{ewoc-irc} in the context of general \glspl{ewoc}) and can therefore be computed in perturbation theory.
    %
    In particular, the use of energy weighting ensures that the EEC is unchanged under the addition of zero-energy radiation (infra-red safety), and the energy weighting together with the dependence of the delta function solely on the angles of particles ensures that it is unchanged under exactly collinear splittings of final-state particles (collinear safety).
    %
    Furthermore, the energy weighting in the integrand also suppresses non-perturbative corrections arising from soft physics to perturbative calculations of the EEC.
    %
    The IRC robustness of the EEC is one it is useful tool for the extraction of the strong coupling constant from particle collision data.
}

@@

Therefore, rather than focusing on the traditional projections of ENCs, in this section we focus on new parametrizations of higher point energy correlators, introduced in \Reff{Alipour-fard:2024szj}, with a number of improved properties.
%
First, our parametrization of the \gls{penc} depends on the largest distance \(R_1\) to a ``special'' particle $s$ in a set of $N$ particles, suitably averaged over all choices for $s$;
%
this yields simpler phase space restrictions than the traditional parametrization for the \gls{penc} in terms of the largest pairwise angle~\cite{Chen:2020vvp}.
%
Second, when considering more differential information, our parametrization of resolved ENCs (\gls{renc}s) employs non-redundant polar coordinates centered around the special particle, as in \Fig{cartoon}, yielding intuitive visualizations of jet substructure.
%
This differs from the traditional approach, which uses over-complete information from the set of all pairwise distances and neglects information about the relative orientation of particles.
%
Third, our parametrization offers dramatic improvements in computational performance, which is essential for their use in experimental analyses.
%
Finally, we anticipate that these conceptual and computational improvements will yield simpler theoretical calculations.
%
Our implementation for computing the \glspl{penc} of \Eq{new_def} be found on GitHub as an update to \textsc{FastEEC}~\cite{Budhraja:2024xiq}, and of our \glspl{penc} and \gls{renc}s at \texttt{ResolvedEnergyCorrelators}~\cite{github:RENC}.

@@

\indent
A major strength of the parametrization of \Eq{new_def} it its dramatically improved computational efficiency relative to older \gls{penc} paramterization.
%
The computational efficiency of our parametrization is more clear in the expression for (normalized) cumulative distribution:
%%%
\begin{align} \label{eq:new_cml}
  \Sigma_N(R_1)
  \!
  &=
  \!
  \frac{1}{\sigma}
  \int_0^{R_1}
  \!\!
  \dd R_1'
  \,
  \frac{\dd \sigma_N}{\dd R_1'}
   % \notag \\
  =
  % \!
  \biggl\langle \sum_{\izero} z_{\izero} [z_{\rm disk}(\izero,R_1)]^{N-1}
  \biggr\rangle
  ,
\end{align}
%%%
where $z_{\rm disk}(\izero,R_1)$ denotes the total transverse momentum fraction of all particles within a radius $R_1$ of the special particle \izero{}.
%
Notably, the simple form of \Eq{new_cml} holds even for non-integer $N$.
%
A practical way to evaluate \Eq{new_cml} (and \Eq{new_def} after differentiation) is, for each $\izero$, to first sort all particles by their distance with respect to $\izero$, and then to compute $\Sigma_N(R_1)$ by beginning with $R_1=0$ and then increasing it.
%
Sorting particles by their distance to \izero{} takes $M\ln M$ time, and the remaining sum over \izero{} scales with $M$, resulting in an overall computation time scaling as $M^2 \ln M$.
%
This computational speed up is especially interesting for heavy-ion collisions where $M$ is typically very large.

>>>>

Importantly, we have left the sum on \(i,j\) a little bit vague.
%
What are we summing over?
%
The answer is that if we use \(i\) and \(j\) to index particles, so that the \gls{ewoc} captures pair-wise particle correlations, we quickly run into trouble.
%
In particular, collinear safety requires that particle-level \glspl{ewoc} can \textit{only} encode angular correlations within a jet:
%
\(\mathcal{O}_{ij} = \theta_{ij}\) or \(\mathcal{O}_{ij} = \Delta R_{ij}\).
%
In order to obtain \glslink{irc-safety}{IRC safe} observable, we must either study only \textit{angular correlations} or \textit{give up on particle-level correlations} in favor of correlations between collective degrees of freedom.%
\footnote{
    We use subjets as proxies for collective degrees of freedom so that our energy-weighted correlations are not fully \glslink{continuity}{continuous} (the number of subjets in a jet, and the discrete yes/no choice of whether a particle belongs to a given subjet, are both discontinuous).
    %
    The only fully continuous substructure observables of this chapter are the EEC and its multi-particle generalizations.
    %
    Nonetheless, the study of \glspl{ewoc} will lead to results which are \textit{more} robust to hadronization.
}

@@

Among jet substructure observables, the Energy-Energy Correlator (EEC) \cite{Basham:1978bw,Basham:1978zq,Basham:1979gh} is particularly noteworthy:
%
it is conceptually simple, insensitive to low-energy radiation, and naturally separates physics at different scales.
%
Physically, the EEC measures the mean correlation in the energy passing through any two detectors separated by an angle \(\Delta\), averaged over many particle collisions:
%
\(
    \text{EEC}(\Delta)
    \propto
    \left\langle
        E_\text{det.\.1} \,\,\, E_\text{det.\.2}
    \right\rangle
\).%
\footnote{
    More concretely, this means that the contribution of each particle in an event to the EEC is weighted by the particle's energy,
    %
    as in \Eq{intro_ewoc_def} and, shorty, \Def{eec}.
    %
    This energy weighting mitigates the effects of the low-energy physics of QCD that complicates the theoretical description of other jet substructure observables.
}
%
From the birth of the EEC to its use in recent collider studies, the simplicity of the EEC has enabled highly accurate perturbative computations \cite{Clay:1995sd,Glover:1994vz,Kramer:1996qr,DelDuca:2016csb,Gituliar:2017umx,Dixon:2018tpg,Dixon:2018qgp,Henn:2019gkr,Luo:2019nig,Gao:2020vyx,Neill:2022lqx,Lee:2023npz,Kramer:1995qh,deFlorian:2004mp,Banfi:2002vw,Tulipant:2017ybb,Moult:2018jzp,Korchemsky:2019nzm,Dixon:2019uzg,Gao:2019ojf,Luo:2019hmp,Luo:2019bmw,Moult:2019vou,Li:2020bub,Ebert:2020sfi,Li:2021txc,Duhr:2022yyp,Chen:2023zlx,Gao:2023ivm} and has facilitated measurements of the strong coupling constant \(\alpha_s\) \cite{Martin:1986uq,DELPHI:1990sof,SLD:1994yoe,ATLAS:2015yaa,ATLAS:2017qir,dEnterria:2018cye,Kardos:2018kqj,Ali:2020ksn,dEnterria:2022hzv,ATLAS:2023tgo} including the most precise jet substructure determinations of \(\alpha_s\) to date \cite{CMS:2024mlf}.
%
The EEC has been investigated as a probe of jets traveling through the quark-gluon plasma (QGP) produced in heavy-ion collisions \cite{Lokhtin:2004tx,Lokhtin:2006dp,Andres:2022ovj,Barata:2023zqg,Andres:2023xwr,Yang:2023dwc,Barata:2023vnl,Barata:2023bhh,Barata:2024nqo} and as a tool in studies of nuclear tomography \cite{Karapetyan:2019fst,Liu:2022wop,Liu:2023aqb,Kang:2023gvg,Cao:2023oef}.
%
The study of the EEC has also shed light on the structure of non-perturbative QCD \cite{Nason:1995np,Korchemsky:1997sy,Korchemsky:1999kt,Dokshitzer:1999sh,Chen:2020vvp,Jaarsma:2023ell,Schindler:2023cww,Lee:2024esz}
and of quantum field theory in general \cite{Richards:1983sr,Sveshnikov:1995vi,Hofman:2008ar,Hatta:2012kn,Belitsky:2013bja,Belitsky:2013ofa,Belitsky:2013xxa,Belitsky:2014zha,Korchemsky:2015ssa,Goncalves:2014ffa,Farnsworth:2015hum,Hofman:2016awc,Kravchuk:2018htv,Kologlu:2019mfz,Chang:2020qpj,Korchemsky:2021htm,Caron-Huot:2022eqs,Chen:2023wah,Chicherin:2023gxt,Chicherin:2024ifn}.
%
It also been put forward as a way to search for new physics in hadronic decays of electroweak bosons~\cite{Ricci:2022htc}.
%
Finally, while most jet substructure observables are dominated by contributions from two-particle correlations, the EEC can be naturally extended to multi-point energy correlators which can characterize correlations between an arbitrary number of particles.
%
Recent theoretical work suggests that multi-point generalizations of the EEC may be used to extract the mass of the top quark at the LHC \cite{Procura:2022fid,Holguin:2022epo,Holguin:2023bjf,Pathak:2023tmy,Xiao:2024rol,Holguin:2024tkz}.

@@

\remark{}{
    The EEC is \glslink{irc-safety}{IRC safe} (as we discuss further in \Sec{ewoc-irc} in the context of general \glspl{ewoc}), and can be computed in \gls{pqcd} (see Lore~\ref{lore:irc-safety-lore}).
    %
    In particular, the use of energy weighting ensures that the EEC is unchanged under the addition of zero-energy radiation (infra-red safety), and the energy weighting together with the dependence of the delta function solely on the angles of particles ensures that it is unchanged under exactly collinear splittings of final-state particles (collinear safety).
    %
    Furthermore, the energy weighting in the integrand also suppresses non-perturbative corrections arising from soft physics to perturbative calculations of the EEC.
    %
    The IRC robustness of the EEC is one it is useful tool for the extraction of the strong coupling constant from particle collision data.
}

@@

Therefore, rather than focusing on the traditional projections of ENCs, in this section we focus on new parametrizations of higher point energy correlators, introduced in \Reff{Alipour-fard:2024szj}, with a number of improved properties.
%
First, our parametrization of the \gls{penc} depends on the largest distance \(R_1\) to a ``special'' particle $s$ in a set of $N$ particles, suitably averaged over all choices for $s$;
%
this yields simpler phase space restrictions than the traditional parametrization for the \gls{penc} in terms of the largest pairwise angle~\cite{Chen:2020vvp}.
%
Second, when considering more differential information, our parametrization of resolved ENCs (\gls{renc}s) employs non-redundant polar coordinates centered around the special particle, as in \Fig{cartoon}, yielding intuitive visualizations of jet substructure.
%
This differs from the traditional approach, which uses over-complete information from the set of all pairwise distances and neglects information about the relative orientation of particles.
%
Third, our parametrization offers dramatic improvements in computational performance, which is essential for their use in experimental analyses.
%
Finally, we anticipate that these conceptual and computational improvements will yield simpler theoretical calculations.
%
Our implementation for computing the \glspl{penc} of \Eq{new_def} can be found on GitHub as an update to \textsc{FastEEC}~\cite{Budhraja:2024xiq}, and of our \glspl{penc} and \gls{renc}s at \texttt{ResolvedEnergyCorrelators}~\cite{github:RENC}.

@@

\indent
A major strength of the parametrization of \Eq{new_def} is its dramatically improved computational efficiency relative to older \gls{penc} paramterization.
%
The computational efficiency of our parametrization is more clear in the expression for (normalized) cumulative distribution:
%%%
\begin{align} \label{eq:new_cml}
  \Sigma_N(R_1)
  \!
  &=
  \!
  \frac{1}{\sigma}
  \int_0^{R_1}
  \!\!
  \dd R_1'
  \,
  \frac{\dd \sigma_N}{\dd R_1'}
   % \notag \\
  =
  % \!
  \biggl\langle \sum_{\izero} z_{\izero} [z_{\rm disk}(\izero,R_1)]^{N-1}
  \biggr\rangle
  ,
\end{align}
%%%
where $z_{\rm disk}(\izero,R_1)$ denotes the total transverse momentum fraction of all particles within a radius $R_1$ of the special particle \izero{}.
%
Notably, the simple form of \Eq{new_cml} holds even for non-integer $N$.
%
A practical way to evaluate \Eq{new_cml} (and \Eq{new_def} after differentiation) is, for each $\izero$, to first sort all particles by their distance with respect to $\izero$, and then to compute $\Sigma_N(R_1)$ by beginning with $R_1=0$ and then increasing it.
%
Sorting particles by their distance to \izero{} takes $M\ln M$ time, and the remaining sum over \izero{} scales with $M$, resulting in an overall computation time scaling as $M^2 \ln M$.
%
This computational speed up is especially interesting for heavy-ion collisions where $M$ is typically very large.

### 2025-05-05 frontmatter/abstract.tex
Pre-correction version of thesis

\begin{Abstract}
\begin{changemargin}{0.5cm}{0.5cm}
This thesis presents the author's work in developing probes of the inner structure of jets in high-energy particle collisions.
%
We begin by introducing QCD and the scattering of partons (quarks and gluons), discussing jets as theoretical and experimental proxies for partonic physics, and presenting the partonic cascade model of jet formation and jet substructure.
%
Noting the ubiquitous presence of low-energy pollution in particle collision events, in the forms of hadronization, detector effects, the underlying event (UE), and pileup (PU), we then move towards the modern research area of developing pollution-insensitive probes of the jet substructure.
%
Pollution-insensitive features of jet substructure are usually accessed either through jet grooming or energy-weighted correlation functions.
%
We present the basics of the modern theory of jet grooming as well as the work of the author in developing the \textsc{Piranha} paradigm for \textit{continuous} jet grooming, introduced by the author in \Reff{Alipour-Fard:2023prj}, and explore the formal and phenomenological benefits of continuous grooming techniques as pollution-insensitive probes of jet substructure.
%
We also introduce the basics of the simplest energy-weighted correlation function -- the energy-energy correlator (EEC), which probes angular correlations between particle pairs -- and discuss the efficient and intuitive higher-point analogues introduced by the author in \Reff{Alipour-fard:2024szj}, which provide computationally-realistic, pollution-insensitive probes of angular many-body correlations in QCD jets.
%
Finally, we exposit the generic theory of energy-weighted observable correlations (EWOCs), introduced by the author in \Reff{Alipour-fard:2025dvp}, which utilize the energy weighting of the EEC to provide pollution-insensitive probes of non-angular correlations within jets.
\end{changemargin}
\end{Abstract}

>>>>

\begin{Abstract}
\begin{changemargin}{0.5cm}{0.5cm}
This thesis presents the author's work in developing probes of the inner structure of jets in high-energy particle collisions.
%
We begin by introducing QCD and the scattering of partons (quarks and gluons), discussing jets as theoretical and experimental proxies for partonic physics, and presenting the partonic cascade model of jet formation and jet substructure.
%
Noting the ubiquitous presence of low-energy pollution in particle collision events, in the forms of hadronization, detector effects, the underlying event (UE), and pileup (PU), we then move towards the modern research area of developing pollution-insensitive probes of jet substructure.
%
Pollution-insensitive features of jet substructure are often accessed theoretically either through jet grooming or energy-weighted correlation functions.
%
We present the basics of the modern theory of jet grooming as well as the work of the author in developing the \textsc{Piranha} paradigm for \textit{continuous} jet grooming, introduced by the author in \Reff{Alipour-Fard:2023prj}, and explore the formal and phenomenological benefits of continuous grooming techniques as pollution-insensitive probes of jet substructure.
%
We also introduce the basics of the simplest energy-weighted correlation function -- the energy-energy correlator (EEC), which probes angular correlations between particle pairs -- and discuss its many-particle analogues.
%
We focus on the efficient and visually intuitive projected and resolved energy correlators introduced by the author in \Reff{Alipour-fard:2024szj}, which provide computationally-realistic, pollution-insensitive probes of angular many-body correlations in QCD jets.
%
Finally, we exposit the generic theory of energy-weighted observable correlations (EWOCs), introduced by the author in \Reff{Alipour-fard:2025dvp}, which utilizes the energy weighting of the EEC to provide pollution-insensitive probes of non-angular correlations within jets.
\end{changemargin}
\end{Abstract}
